Baseline centering problem

### 2018-12-19 paper/tangotd.tex

\paragraph{Better variant?}
I think
everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem; still, it looks suspiciously like $TD(1)$ except maybe for the
$O(1/(1-\gamma))$ variance). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.

@@

On-policy, working with $\gamma\to 1$ is quite easy: indeed $Z(s_1,s_2)$ is
a density with respect to $\mu(s_2)$ so that the asymptotics when
$\gamma\to 1$ is $Z(s_1,s_2)=\frac{1}{1-\gamma}+Z_0(s_1,s_2)$, and we can
directly learn $Z_0$ by inserting $\frac{1}{1-\gamma}+Z_0$ in the middle
Bellman equation. This yields (omitting the expectations)
\begin{multline}
Z_0(s_1,s_2)\gets (1+\eta)Z_0(s_1,s_2)-\eta
Z_0(s_1,s_3)Z_0(s_3,s_2)+\\\eta\gamma Z_0(s_1,s_3)Z_0(s'_3,s_2)-\eta
Z_0(s_1,s_3)+\frac{\eta}{1-\gamma}(\gamma Z_0(s'_3,s_2)-Z_0(s_3,s_2))
\end{multline}
TODO CHECK which we cannot directly apply with $\gamma=1$...

>>>>

\paragraph{Better variant?}
Everything works equally well if we just learn a kernel $Z(s_1,s_2)$ such
that
\begin{equation}
\transp{\1_{s_1}}\Delta^{-1}\psi\approx \E_{s_2\sim \rho} \,Z(s_1,s_2)
\psi(s_2)
\end{equation}
and this is still linear in $\psi$. This represents the inverse Laplace
operator by
\begin{equation}
Z=\sum_{s_1,s_2}\rho(s_2)  Z(s_1,s_2)\1_{s_1}\transp{\1_{s_2}}
\end{equation}
In particular the value function is
\begin{equation}
V(s_1)=\E_{s_2\sim \rho} \, Z(s_1,s_2) R(s_2)
\end{equation}
Downside: this representation of the value function does not allow for first
computing a representation of $R$ as $\E_{s_2} R(s_2)f(s_2)$ online, then taking a dot product with
$\green$. Here we would have to learn an auxiliary network to imitate
$\E_{s_2\sim \rho} Z(s_1,s_2) R(s_2)$ (but this is a supervised learning
problem; still, it looks suspiciously like $TD(1)$ except maybe for the
$O(1/(1-\gamma))$ variance). Alternatively, policy gradient can just directly be obtained
via
\begin{equation}
\partial \pi(a|s)\E_{s'|s,a} \E_{s_2\sim \rho} \, Z(s',s_2)R(s_2)
\end{equation}
(this works even if $\rho$ is \emph{not} the invariant distribution). We
can apply this each time a new reward is observed at some point $s_2$. We
can even set another arbitrary distribution on $s_1$: this policy
gradient will maximize
the reward obtained from a starting point from that distribution.

@@

On-policy, working with $\gamma\to 1$ is quite easy: indeed $Z(s_1,s_2)$ is
a density with respect to $\mu(s_2)$ so that the asymptotics when
$\gamma\to 1$ is $Z(s_1,s_2)=\frac{1}{1-\gamma}+Z_0(s_1,s_2)$, and we can
directly learn $Z_0$ by inserting $\frac{1}{1-\gamma}+Z_0$ in the middle
Bellman equation. This yields (omitting the expectations)
\begin{multline}
Z_0(s_1,s_2)\gets (1+\eta)Z_0(s_1,s_2)-\eta
Z_0(s_1,s_3)(Z_0(s_3,s_2)-\gamma Z_0(s'_3,s_2))\\-\eta
Z_0(s_1,s_3)+\frac{\eta}{1-\gamma}(\gamma Z_0(s'_3,s_2)-Z_0(s_3,s_2))
\end{multline}
TODO CHECK which we cannot directly apply with $\gamma=1$...
However ON-POLICY $s_3$ and
$s'_3$ have the same law, so in expectation the last term is $-\eta
\E_{s_3} Z_0(s_3,s_2)$. This yields the more symmetric
\begin{multline}
Z_0(s_1,s_2)\gets (1+\eta)Z_0(s_1,s_2)-\eta
Z_0(s_1,s_3)\left(Z_0(s_3,s_2)-\gamma Z_0(s'_3,s_2)\right)\\-\eta
Z_0(s_1,s_3)-\eta Z_0(s_3,s_2)
\end{multline}
on-policy, for any $\gamma\leq 1$.
